---
title: "esm206_lab7"
author: "Ali Martin"
date: "11/8/2021"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message  = FALSE, warning = FALSE)

#Attach packages
library(tidyverse)
library(palmerpenguins)
library(GGally)
library(broom)
library(kableExtra)
library(modelsummary)
library(stargazer)
```



## Intro to GGpairs

Use GGally:: ggpairs() function to create a pairs plot

```{r}
ggpairs(penguins)
```

```{r}
penguins %>% 
  select(species, bill_length_mm: body_mass_g) %>% # the single colon :shows selected columns from bill length all the way to body mass
ggpairs(aes(color= species)) #color to change based on species

ggplot(data = penguins, aes(x = bill_depth_mm, y = body_mass_g, group = species)) +
  geom_point(aes(color = species)) +
  geom_smooth(method = "lm") # add regression line
#there is a positive linear relationship between bill depth and body mass
#penguin species is the important variable to group by that shows R that there is a positive relationship
```

## Multiple Linear regression- penguins

Use multiple linear regression to try and understand the relationship bnetween multiple predictor variables and a single continuous outcome variable (body mass).

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm + bill_depth_mm, data = penguins) #write in formula notation: dependent variable on left side (as a function) of the predictor variables on the right side

#return just the coefficients:
penguin_lm

#for this model, we can interpret the coefficient of flipper_length_mm, 51.54 as on average we would expect a 1 mm increase in flipper length to be associated with a 51.54 g increase in body mass (if everything else is constant). 
```

We would expect, on avergae, a 51.54 g increase in penguiin body mass for each 1 mm increase in flipper legnth, if everything else is constant. 

We would expect, on avergae, a 22.63 g increase in body mass for each 1 mm increase in bill depth. 

Use `summary()` function to get more information:

```{r}
summary(penguin_lm)
```
The adjusted R squared value of 0.7596 means there is 76% of variance in penguin body mass is explained by this model which means there is 24% of body mass that isn't explained. There is no good or bad R squared value. 

### Use `broom` functions tidy and glance

Use broom ::tidy to get the coefficient summary in a data frame.
Use broom:: glance to get the model summary in a data frame. 

```{r}
penguin_lm_tidy <- tidy(penguin_lm)
penguin_lm_tidy
```

From the estimate column in penguin lm to pull the third column (bill depth)
Try using the in-line code to refer to (print) the bill depth coefficient
r spedific data frame $estimatecolumn [# of element]
`r penguin_lm_tidy$estimate[3]`

Glance can be used to look at model outputs

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```
# Instead of using just bill depth and flipper length, variables that are important to comapre body mass in penguins would be species and sex 

## Try another model
We have thought really hard, done a literature review, consulted with experts, deeply explored our data and decided that the three most critical variables to include in this model (with body mass as the dependent variable) are:

-flipper length
-species
-sex

Task: create a multiple linear regression model using those 3 predictor variables

Then take a look at the model outputs and interpret the coefficients with a neighbor

```{r}
penguin_lm_2 <- lm(body_mass_g ~ flipper_length_mm + species + sex , data = penguins) 

penguin_lm_2
```

Reference species is Adelie (because it does not show up), reference sex is female 

Equation: mass(g) = -365.82 + 20.02(flip(mm)) - 87.63(chinstrap) +836.26(gentoo) +530.38 (male)

we are interpreting these coefficients only in reference to the reference variables

If everything else is the same, on average we expect a gentoo penguin to weigh 836.26 grams more than an Adelie penguin.

Similarly,if everything is the same,  we expect chinstrap penguins to be smaller than Adelie's on average of 87.63 grams

If everything else is the same, we expect a male penguin to weigh 530.38 g more than a female penguin.

For a female chinstrap penguin with flip length of 180 g our equation looks like this (plug in the values):
mass(g) = -365.82 + 20.02 (180) - 87.63(1) +836.26(0) + 530.38 (0) <- reference equation above


### Use broom:: augment() to see predicted values 

```{r}
augment(penguin_lm_2)
```
## Use plot() to look at model diagnostics

```{r}
plot(penguin_lm_2)
```

-Conceptual understanding and expertise allows me to responsibly determine that penguin species, sex and flipper length are reasonable predictor variables (make sure you are thinking ahrd about variables and different variable combinations)
-Diagnostic plots all check out for nomality of residuals, homoscedasticity, and no worrisome outliers (<< last one is not an assumption)

## Make output tables

Using model summary:

```{r}
modelsummary(penguin_lm_2)
```

Using stargazer

```{r, results = "asis"}
stargazer(penguin_lm_2, type = "html")
```

